Large-scale neural circuits on an FPGA

### 2015-05-14 docs/paper/paper.tex
added first two sections
-finished NEF description
- finished intro
- finished neuromorphic section

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{authblk}
\usepackage{amsmath}

\newcommand{\design}{Nengo-RT}

@@

\author[1]{Murphy Berzish}
\author[1,2,3]{Chris Eliasmith}
\author[1,2]{Bryan Tripp}
\affil[1]{Centre for Theoretical Neuroscience, University of Waterloo}
\affil[2]{Dept. of Systems Design Engineering, University of Waterloo}
\affil[3]{Dept. of Philosophy, University of Waterloo}
\maketitle

\abstract{abstract goes here - bryan???}

@@

\subsection{Motivation}

- potential applications

\subsection{State of the Art (Neuromorphic Hardware)}

- from my other paper for mind collection

@@

\subsection{The Neural Engineering Framework}

@@

Given an encoding operation, it is natural to define a \textit{decoding} operation,
in order to characterize the information processing characteristics
of the system (in this case an ensemble of neurons). To decode a continuous
estimate of the input from neural activity, the NEF focuses on the
postsynpatic filtered activity generated by the reception of a spike
at a synapse. This activity is taken to be a linear filter applied
to the spiking activity:
\[
a_{i}(\mathbf{x})=\sum_{j}h_{i}(t)*\delta_{i}(t-t_{j}(\mathbf{x}))
\]
where $h_{i}(t)$ is the synaptic response function (usually a decaying
exponential) with a time constant, $\tau_{PSC}$, determined by neurotransmitter
type at the synapse, '$*$' is the convolution operator, and $\delta_{i}(t-t_{j}(\mathbf{x}))$
is the spike train produced by neuron $i$ in response to input $\mathbf{x}$,
with spike times indexed by $j$. Having defined this continuous variable,
which is equal to the spike rate as $\tau_{PSC}\rightarrow\infty$,
we can specify a decoding operation for estimating the input $\mathbf{x}$.

For reasons that will be apparent in a moment, we use a \textit{linear}
decoder:

@@

Any optimization method can be used to find these decoders. The simplest
is to use standard least-squares optimization:
\begin{equation}
\underset{\mathbf{d}_{i}}{arg\, min}\int[\mathbf{x}-\sum_{i}^{N}a_{i}(\mathbf{x)}\mathbf{d}_{i}]^{2}d\mathbf{x}\label{eq:error}
\end{equation}
where $\mathbf{d}_{i}$ are the decoding vectors over which this error
is minimized and the integral is over all $\mathbf{x}$ values. It
has been shown that linear decoders are sufficient to decode \textasciitilde{}95\%
of the information available in a single spike train generated by
a stimulus \cite{Rieke1997b}. Furthermore, as the number of neurons
$N$ increases, the mean squared error decreases as $1/N$. The NEF
decoding process is depicted in Figure \ref{fig:NEF-decoding}, where
the optimal linear decoders have been found and used for twenty neurons.
Temporal decoding is also linear, as described, and is performed using
the postsynaptic, current-based filters (see section \ref{sub:Principle-3--}
for further discussion).

While linear decoders are useful for visualizing the information encoded
within the activity of a group of neurons, they also provide a direct
way to solve for connection weights between groups of neurons. This
is a key advantage of the NEF: rather than using a learning rule to
optimize over the entire space of all connection weights, we instead
solve the simpler problem of optimizing over the space of linear decoders,
and then use that result to solve for the connection weights. Importantly,
given the characterization of neural activity with preferred direction
vectors, there is no difference between using this smaller space and
using the equivalent full connection matrix. Interestingly, this general
technique has started to be applied in broader domains; see \cite{tapson2013}
for an overview.

For example, if a connection between neural groups is meant to compute
the identity function $\mathbf{y}=\mathbf{x}$ (where $\mathbf{y}$
is the vector space represented by the second population B and $\mathbf{x}$
is the vector space represented by the first population A), the connections
between individual neurons are given by 
\begin{equation}
\omega_{ij}=\alpha_{j}\mathbf{d}_{i}\mathbf{e}_{j}\label{eq:weights}
\end{equation}
where $i$ indexes the neurons in group A and $j$ indexes the neurons
in B. Behavior of this network is shown in Figure \ref{fig:Using-NEF-weights}a.

While the least-squares method for optimization we use here is not
biologically plausible on its own, we have also shown that biologically
realistic learning rules will converge on a similar solution \cite{MacNeil2011a}.
These realistic rules are, however, several orders of magnitude more
computationally expensive.

@@

Connections between groups of neurons can also compute functions other
than the identity function. That is instead of \textbf{$\mathbf{y}=\mathbf{x}$}
we can do \textbf{$\mathbf{y}=f(\mathbf{x})$}. We do this by finding
decoders \textbf{$\mathbf{d}_{i}^{f}$ }for the particular function\textbf{
$f(\mathbf{x})$} by substituting \textbf{$f(\mathbf{x})$ }for $\mathbf{x}$
in equation (\ref{eq:error}).

@@

The connection weights can then be computed using (\ref{eq:weights}).
For the special case of linear functions\textbf{ $\mathbf{y}=\mathbf{L}\mathbf{x}$},
rather than solving for a new decoder, we can simply put $\mathbf{L}$
directly into the weight equation itself, resulting in $\omega_{ij}=\alpha_{j}\mathbf{d}_{i}\mathbf{L}\mathbf{e}_{j}$.
Combining these two approaches, the neural connection weights needed
to approximate the function\textbf{ $\mathbf{y=L}f(\mathbf{x\textrm{)}}$}
are:
\begin{equation}
\omega_{ij}=\alpha_{j}\mathbf{d}_{i}^{f}\mathbf{L}\mathbf{e}_{j}\label{eq:weights-general}
\end{equation}
Figure \ref{fig:Using-NEF-weights}b shows the computation of the
element-wise square ($f(\mathbf{x})=[x_{1}^{2},x_{2}^{2}]$).

@@

While the first two principles are sufficient to build neural approximations
of any desired function of the vector $\mathbf{x}$, the NEF also
provides a method for computing functions of the form \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}f(\mathbf{x},\mathbf{u})$},
where $\mathbf{u}$ is the input from some other population. We do
this by exploiting the fact that neurons do not simply accept input
as spikes. Rather, when a spike is transmitted from one neuron to
another, the actual current that flows into the second neuron is a
low-pass-filtered version of that spike. In particular, the post-synaptic
current is well-approximated by $h(t)=u(t)e^{-t/\tau}$, where $u(t)$
is the step function and $\tau$ is the time constant of the neurotransmitter
used. This time constant varies throughout the brain, e.g., from 2-5
ms (AMPA; \cite{jonas1993quantal}) up to \textasciitilde{}100ms (NMDA;
\cite{sah1990properties}). The effect of this filter is that instead
of a connection computing the function $\mathbf{y}(t)=f(\mathbf{x}(t))$
it will compute $\mathbf{y}(t)=f(\mathbf{x}(t))*h(t)$ (or, in the
Laplace domain, $\mathbf{Y}(s)=\mathbf{F}(s)H(s)$).

It turns out that this implicit low-pass filter can be used to generate
neural models that compute abitrary dynamics. Given a neural population
representing $\mathbf{x}$, an input $\mathbf{u}(t)$, and a feedback
connection from $\mathbf{x}$ back to itself computing $g(\mathbf{x}(t))$,
we note that in the Laplace domain we get $\mathbf{X}(s)=(\mathbf{G}(s)+\mathbf{U}(s))H(s)$.
Since the Laplace transform of $h(t)=u(t)e^{-t/\tau}$ is $H(s)=1/(1+s\tau)$,
we can rearrange this to get $s\mathbf{X}(s)=(\mathbf{G}(s)-\mathbf{X}(s))/\tau+\mathbf{U}(s)/\tau$.
Converting back to the time domain, \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}\frac{g(\mathbf{x}(t))-\mathbf{x}(t)}{\tau}+\frac{\mathbf{u}(t)}{\tau}$}.
Thus, if we desire the dynamics \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}f(\mathbf{x}(t))+\mathbf{u}(t)$},
we introduce a feedback connection that uses the previous two NEF
principles to find connection weights that compute $g(\mathbf{x}(t))=\tau f(\mathbf{x})+\mathbf{x}$
and we scale the input $\mathbf{u}(t)$ by $\tau$. For arbitrary
dynamics of the form \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}f(\mathbf{x},\mathbf{u})$},
we do a change of variables $\mathbf{x}^{\prime}=\langle\mathbf{x},\mathbf{u}\rangle$.

This exploitation of the inherent first-order low-pass filter found
in synaptic connections allows for the implementation of a very wide
variety of systems, including oscillators, integrators, and arbitrary
attractor networks \cite{Eliasmith2005p}. Figure \ref{fig:Two-dynamical-systems}a
shows a single neural population with a recurrent feedback connection
computing the standard linear oscillator \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}[-x_{2},-x_{1}]$}.
Here, the post-synapic time constant $\tau$ is 100ms, but the period
of oscillation is $2\pi$ (\textasciitilde{}6.28 seconds). Importantly,
we achieve different periods of oscillation without changing the neural
property $\tau$; rather, we change the function being computed by
the feedback connection, resulting in a different set of connection
weights. For example, we could compute \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}[-x_{2}/2,-x_{1}/2]$},
giving a period of $\pi$, instead. This same method works for nonlinear
functions as well; in Figure \ref{fig:Two-dynamical-systems}b we
show the classic Lorenz attractor \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}[10(x_{2}-x_{1}),x_{1}(28-x_{3})-x_{2},x_{1}x_{2}-\frac{8}{3}x_{3}]$}.
This approach allows for the construction of neural models that correspond
to a very large family of functions, including those typically employed
by modern control theory and dynamic systems theory.

It should be noted that networks constructed with this approach are
fundamentally different from Echo State Networks \cite{jaeger2001}
and Liquid State Machines \cite{Maass2002}. In both of those approaches,
the recurrent connections are randomly chosen, and then a linear combination
of the outputs are found that compute the desired function. With the
NEF, we also find a linear combination of the outputs ($\mathbf{d}_{i}$),
but we solve for the ideal recurrent connection weights to achieve
the desired dynamics. This makes the NEF much more efficient and capable
of computing a much wider range of dynamics.

@@

The principles of the NEF have been incorporated into the Nengo neural simulator \cite{Bekolay2014}.  This simulator

- mention spaun

These three principles are sufficient to implement all of our neural
models. However, to simplify the process of constructing these models,
we have developed an open-source software package known as Nengo (Neural
ENGineering Objects) that creates and runs these models. Models can
be built in Nengo using a drag-and-drop graphical user interface or
specified using the Python scripting language. Full details and documentation
can be found online at http://nengo.ca, and in other publications
\cite{Stewart2009l}.

For example, to create the model shown in Figure \ref{fig:Using-NEF-weights}a,
we use the following script:
\begin{verbatim}
net = nef.Network("Identity Function")
net.make("A", neurons=20, dimensions=1) 
net.make("B", neurons=20, dimensions=1)
net.connect("A", "B") 
\end{verbatim}

For Figure \ref{fig:Using-NEF-weights}b we need to compute the element-wise
square. This is specified in a Python function as follows:

\begin{verbatim}
net = nef.Network("Element-wise Square")
net.make("A", neurons=20, dimensions=1) 
net.make("B", neurons=20, dimensions=1)

def square(x):
    return x{[}0{]}{*}x{[}0{]}, x{[}1{]}{*}x{[}1{]}

net.connect("A", "B", func=square) 
\end{verbatim}

Nengo will automatically solve for the connection weights that will
best approximate the provided function.

Nengo also provides an interactive interface for displaying the results
of a simulation while it is running, allowing for real-time interaction
with a running model. This interface allows the generated plots to
be exported, and was used to produce the previous figures in this
paper (i.e. all figures in section \ref{sec:The-Neural-Engineering}).
Furthermore, Nengo scales up to our largest models: the 2.5 million
neuron Spaun model is run in Nengo, and can be downloaded at http://models.nengo.ca/spaun.

- backends

@@

- Bryan???

@@

\design{} is a custom digital hardware design that implements the population-mode NEF in real-time (1~ms timestep).
A block diagram of the design is shown in Figure~\ref{fig:system}.
The encoder units perform the function of encoding (), % FIXME NEF equation
while the population units perform the decoding () % FIXME NEF equation
and write the decoded values to the decoded value buffers
to be used as inputs to other populations or as outputs from the network.
Each population of neurons is assigned to a population unit, which is able to simulate up to 1024 populations
having either one or two encoded inputs each. These inputs are calculated by the encoder units,
of which there are two for a one-dimensional population unit and four for a two-dimensional population unit;
each encoder unit can be associated with a different first-order filter to simulate post-synaptic filtering.
For example, one of the two encoders on a one-dimensional population unit can be programmed with
a small time constant to model fast synapses (e.g. AMPA and GABA$_A$), while the second can be programmed
with a large time constant to model slow synapses (e.g. NMDA and GABA$_B$).
Each population unit performs time-multiplexing and swaps in different sets of decoders
in order to run more populations on the same hardware. The principal components are fixed and must be shared
between all populations on an individual population unit.
Each population unit writes to a fixed set of decoded value buffers, but a shared interconnect
allows the encoder units to perform random accesses to these buffers.
This allows effective all-to-all communication between populations while reducing the number of 
connections that must be made between population units.
A number of decoded value buffers are reserved for external inputs to the network,
which can be sent from the controlling computer system over an Ethernet connection.
The output channels can also be programmed to send decoded values back to the computer over Ethernet.

@@

Nengo supports a more advanced connection called a learning connection,
which mdoifies the decoders of a population according to a particular ``learning rule''
as a function of other represented values. The hardware design does not support this special connection type.
It was decided to implement the most common cases first in order to demonstrate the feasibility of simulating
neural networks in population mode on the hardware, and since many networks do not use learning connections,
support for them was omitted from this version.

>>>>

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{url}
\usepackage{marvosym}

\newcommand{\design}{Nengo\char`_FPGA}  %% Could be xFPGA to specify xilinx? since there might be other fpga backends

@@

% I don't care about order, just assumed you'd want to be 'senior author'
\author[1]{Murphy Berzish}
\author[1,2,3]{Chris Eliasmith}
\author[1,2]{Bryan Tripp}
\affil[1]{Centre for Theoretical Neuroscience, University of Waterloo}
\affil[2]{Dept. of Systems Design Engineering, University of Waterloo}
\affil[3]{Dept. of Philosophy, University of Waterloo}
\maketitle

\begin{abstract}
abstract goes here - bryan???
\end{abstract}

@@

Large-scale neural models have recently become centrepieces of very large international research investments. The controversial European Human Brain Project (HBP) has a central goal of developing a human-scale neural simulation in the next 10 years, and has been granted \EURtm 1 billion to do so.  Similarly, the Obama Brain Initiative (OBI) in the US is investing well over \$1 billion over a similar timeframe in theoretical and experimental neuroscience. Both projects have identified massive increases in computational performance as a critical for achieving their goals.

One approach to solving this challenge is employing large supercomputers, such as the IBM BlueGene.  Both HBP and OBI have exploited this resource to simulation millions or billions of neurons, although not in real time.  However, such resources are available to only a select few researchers.

Another approach is to build specialized hardware -- often called ``neuromorphic'' hardware -- to massively increase the efficiency and speed of neural simulations.  There have been several exciting developments in this regards, both in industry and in academia.  IBM recently announced what will soon be the first commercially available neuromorphic chip \cite{Merolla2014}. As well, several academic projects, such as the SpiNNaker \cite{Khan2008} and Neurogrid \cite{Choudhary2012} projects have made impressive progress in recent years (see Section \ref{sub:neuromorphic} for further discussion).  However, these solutions are not available in large quantities, and hence remain inaccessible to a majority of researchers.

Here, we focus on developing an affordable and accessible solution to the problem of simulating large-scale neural circuits.  Our approach allows anyone with access to off-the-shelf FPGA hardware to simulate models up to about 10 million equivalent neurons with arbitrary connectivity in real time.

Importantly, this level of scaling is dependent on the notion of ``equivalent neurons.'' In Section {\ref{sub:NEF}} we discuss our theoretical approach that allows extremely good approximations of the activity of about a thousand spiking neurons by simulating them as a single population, at a resolution of about 1~ms.  Simulating large neural circuits at this level of abstraction allows for very computationally efficient solutions that preserve function. The main cost of this excellent computational scaling is that these methods do not allow online tuning of the connection weights between individual cells.  However, a large-class of adaptive and non-adaptive massively parallel circuits can be effectively simulated, providing a useful tool for exploring the class of brain-like algorithms.

Indeed, one of the main benefits of our approach is that it focusses both on function and biological plausibility, in contrast to work presented by HBP and OBI  which captures little function \cite{eliasmith2013a}.  This focus remains here, allowing the methods presented to be applied both to characterizing biological cognition and to exploring engineering applications of brain-like algorithms. To this end, we present examples that compute specific mathematical functions, as well as those that simulate early visual processing of time-varying stimuli.

We begin our discussion by describing related neuromorphic approaches in the next section.  We then explain our methods for constructing neural simulations, and demonstrate the equivalence between single cell and population-level simulations.  Next, we discuss the hardware design and show that it is able to efficiently simulate a variety of small test circuits.  We then show that the system scales efficiently to simulating very large models in real time. We conclude by discussing possible improvements to the current design.

\subsection{\label{sub:neuromorphic}Current Neuromorphic Hardware\protect\footnote{This discussion is based on that in \cite{eliasmith2015a}.}}

Perhaps best known neuromorphic hardware is that being developed by IBM.  The architecture itself is called TrueNorth. It includes a digital model of individual neurons, and a method for programming the architecture \cite{Esser2013}.  The purpose of this architecture is to exploit brain-like organization, to permit the simulation of a specific class of algorithms that do not efficiently run on classical computer architectures. IBM researchers argue that the ``low-precision, synthetic, simultaneous, pattern-based metaphor of TrueNorth is a fitting complement to the high-precision, analytical, sequential, logic-based metaphor of today's of von Neumann computers'' \cite[p.~1]{Esser2013}.

TrueNorth has neurons organized into 256 neuron blocks in which each neuron can receive input from 256 axons. To assist with programming this hardware, IBM has introduced the notion of a ``corelet'', which is an abstraction that encapsulates local connectivity in small networks. These act like small programs that can be composed in order to build up more complex functions. To date the demonstrations of the approach have focused on simple, largely feed-forward, standard applications, although they have employed wide range of methods including RBMs, liquid state machines, HMMs, and very simple NEF networks. It should be noted that the proposed chip does not yet exist, and current demonstrations are on detailed simulations of the architecture. However, because it is a digital chip the simulations are highly accurate.

A direct competitor to IBM's approach is the Zeroth neuromorphic chip from Qualcomm. Like IBM, Qualcomm believes that constructing brain-inspired hardware will provide a new paradigm for exploiting the efficiencies of neural computation for the kinds of information processing that brains excel at, but which are extremely challenging for standard von Neumann approaches. The main difference between these 2 approaches is that Qualcomm has committed to allowing online learning to take place on the hardware. Consequently, they announced their processor by demonstrating its application in a reinforcement learning paradigm on a real-world robot. They have released videos of the robot maneuvering in an environment and learning to only visit one kind of stimulus (white boxes; \url{http://www.youtube.com/watch?v=8c1Noq2K96c}). It should again be noted that this is an FPGA simulation of a digital chip which does not yet exist. However the simulation, like IBM's, is highly accurate.  However, it has recently been suggested that Qualcomm is no longer fully supporting this research effort.

Both of these projects are spearheaded by industry.  In the academic sphere, the Spinnaker project at Manchester University has not focused on designing new kinds of chips, but instead focused on using low-power ARM processors at a massive scale to allow large-scale brain simulations \cite{Khan2008}. As a result, the focus has been on designing approaches to routing that allow for the high bandwidth communication that underwrites much of the brain's information processing. Simulations on the Spinnaker hardware typically employ spiking neurons, like IBM and Qualcomm, and occasionally allow for learning \cite{davies2013spike}, like Qualcomm's approach. However, even with low power conventional chips, the energy usage is projected to be higher on the Spinnaker platform per neuron. Nevertheless, Spinnaker boards have been used in a wider variety of larger-scale embodied, and non-embodied applications. These include simulating place cells, path integration, item classification, and simple sensory guided movements \cite{Stewart2013a}.

There are also a number of neuromorphic projects that use analog instead of digital implementations of neurons. Analog approaches tend to be several orders of magnitude more power efficient \cite{hasler2013}, although also more noisy, unreliable, and subject to process variation. These projects include work on the Neurogrid chip at Stanford University \cite{Choudhary2012}, and on a chip at ETH Z\"{u}rich \cite{Corradi2014}. The Neurogrid chip has demonstrated large numbers of simulated neurons, up to 1 million, while the ETH Z\"{u}rich chip allows for online learning. More recently, the Neurogrid chip has been used to control a nonlinear, 6 degree of freedom robotic arm, exhibiting perhaps the most sophisticated information processing from an analog chip to date \cite{Menon14:biorob}.

In addition to the above neuromorphic projects, which are focused on cortical simulation, there have been several specialized neuromorphic chips that mimic the information processing of different perceptual systems. For example, the dynamic vision sensor (DVS) artificial retina developed at ETH Z\"{u}rich performs real-time vision processing that results in a stream of neuron-like spikes \cite{Lichtsteiner2008}. Similarly, an artificial cochlea called AEREAR2 has been developed that generate spikes in response to auditory signals \cite{Li2012}. The development of these and other neuronal sensors make it possible to build fully embodied spiking neuromorphic systems \cite{Galluppi2014}.

More directly relevant to the present paper, there have be several attempts to build neuromorphic systems on FPGAs.  For instance, \cite{cassidy2011design} describe a design with a million ``neurons'', where each neuron is abstracted as an arithmetic logic unit.  In contrast to our design, all neurons are considered identical, there are no synaptic dynamics, and there is no modelling of neural noise. In \cite{li2012a}, synaptic dynamics are included, but the network consists of a total of 256 neurons.  More recently, in \cite{wang2013} a network of over 4 thousand neurons is implemented.  However, the approach we describe here is focussed on simulations of much larger numbers of effective neurons.

@@

\subsection{\label{sub:NEF}The Neural Engineering Framework}

@@

Given an encoding operation, it is natural to define a \textit{decoding} operation,
in order to characterize the information processing characteristics
of the system. To decode a continuous
estimate of the input from neural activity, the NEF focuses on the
postsynaptic filtered activity generated by the reception of a spike
at a synapse. This activity is taken to be a linear filter applied
to the spiking activity:
\[
a_{i}(\mathbf{x})=\sum_{j}h_{i}(t)*\delta_{i}(t-t_{j}(\mathbf{x}))
\]
where $h_{i}(t)$ is the synaptic response function (usually a decaying
exponential) with a time constant, $\tau_{PSC}$, determined by neurotransmitter
type at the synapse, '$*$' is the convolution operator, and $\delta_{i}(t-t_{j}(\mathbf{x}))$
is the spike train produced by neuron $i$ in response to input $\mathbf{x}$,
with spike times indexed by $j$. Having defined this continuous variable,
which is equal to the spike rate as $\tau_{PSC}\rightarrow\infty$,
we can specify a decoding operation for estimating the input $\mathbf{x}$:

@@

Any optimization method can be used to find these decoders. A simple example is least-squares optimization:
\begin{equation}
\underset{\mathbf{d}_{i}}{arg\, min}\int[\mathbf{x}-\sum_{i}^{N}a_{i}(\mathbf{x)}\mathbf{d}_{i}]^{2}d\mathbf{x}\label{eq:error}
\end{equation}
where $\mathbf{d}_{i}$ are the decoding vectors over which this error
is minimized and the integral is over all $\mathbf{x}$ values. As the number of neurons
$N$ increases, the mean squared error decreases as $1/N$.

Note that employing linear decoding allows us to compute connection weights. This
is a key advantage of the NEF: rather than using a learning rule to
optimize over the entire space of all connection weights, we instead
solve the simpler problem of optimizing over the space of linear decoders,
and then use that result to solve for the connection weights. For example, if a connection between neural groups is meant to compute
the identity function $\mathbf{y}=\mathbf{x}$ (where $\mathbf{y}$
is the vector space represented by the second population B and $\mathbf{x}$
is the vector space represented by the first population A), the connections
between individual neurons are given by 
\begin{equation}
\omega_{ij}=\alpha_{j}\mathbf{d}_{i}\mathbf{e}_{j}\label{eq:weights}
\end{equation}
where $i$ indexes the neurons in group A and $j$ indexes the neurons
in B.

While the least-squares method for optimization is not
biologically plausible on its own, we have shown that biologically
realistic learning rules will converge to a similar solution \cite{MacNeil2011a}.
These realistic rules are, however, several orders of magnitude more
computationally expensive.

@@

Principle 1 effectively assumes that we want to compute the identity function. Connections between groups of neurons can also compute functions other
than the identity function. That is, instead of \textbf{$\mathbf{y}=\mathbf{x}$}
we can compute \textbf{$\mathbf{y}=f(\mathbf{x})$}. We do this by finding
decoders \textbf{$\mathbf{d}_{i}^{f}$ }for the particular function\textbf{
$f(\mathbf{x})$} by substituting \textbf{$f(\mathbf{x})$ }for $\mathbf{x}$
in equation (\ref{eq:error}).

@@

The connection weights can then be computed using (\ref{eq:weights}).
For the special case of linear functions\textbf{ $\mathbf{y}=\mathbf{L}\mathbf{x}$},
rather than solving for a new decoder, we can simply put $\mathbf{L}$
directly into the weight equation itself, resulting in $\omega_{ij}=\alpha_{j}\mathbf{d}_{i}\mathbf{L}\mathbf{e}_{j}$.
Combining these two approaches, the neural connection weights needed
to approximate the function\textbf{ $\mathbf{y=L}f(\mathbf{x\textrm{)}}$}
are:
\begin{equation}
\omega_{ij}=\alpha_{j}\mathbf{d}_{i}^{f}\mathbf{L}\mathbf{e}_{j}\label{eq:weights-general}
\end{equation}

@@

While the first two principles are sufficient to build neural approximations
of any desired feedforward function of the vector $\mathbf{x}$, the NEF also
provides a method for computing functions of the form \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}f(\mathbf{x},\mathbf{u})$},
where $\mathbf{u}$ is the input from some other population. The NEF exploits the fact that neurons do not simply accept input
as spikes. Rather, when a spike is transmitted from one neuron to
another, the actual current that flows into the second neuron is a
low-pass-filtered version of that spike. In particular, the post-synaptic
current is well-approximated by $h(t)=u(t)e^{-t/\tau}$, where $u(t)$
is the step function and $\tau$ is the time constant of the neurotransmitter
used. This time constant varies throughout the brain, e.g., from 2-5
ms (AMPA; \cite{jonas1993quantal}) up to $\sim$100ms (NMDA;
\cite{sah1990properties}). The effect of this filter is that instead
of a connection computing the function $\mathbf{y}(t)=f(\mathbf{x}(t))$
it will compute $\mathbf{y}(t)=f(\mathbf{x}(t))*h(t)$ (or, in the
Laplace domain, $\mathbf{Y}(s)=\mathbf{F}(s)H(s)$).

It turns out that this implicit low-pass filter can be used to generate
neural models that compute arbitrary dynamics. Given a neural population
representing $\mathbf{x}$, an input $\mathbf{u}(t)$, and a feedback
connection from $\mathbf{x}$ back to itself computing $g(\mathbf{x}(t))$,
we note that in the Laplace domain we get $\mathbf{X}(s)=(\mathbf{G}(s)+\mathbf{U}(s))H(s)$.
Since the Laplace transform of $h(t)=u(t)e^{-t/\tau}$ is $H(s)=1/(1+s\tau)$,
we can rearrange this to get $s\mathbf{X}(s)=(\mathbf{G}(s)-\mathbf{X}(s))/\tau+\mathbf{U}(s)/\tau$.
Converting back to the time domain, \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}\frac{g(\mathbf{x}(t))-\mathbf{x}(t)}{\tau}+\frac{\mathbf{u}(t)}{\tau}$}.
Thus, if we desire the dynamics \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}f(\mathbf{x}(t))+\mathbf{u}(t)$},
we introduce a feedback connection that uses the previous two NEF
principles to find connection weights that compute $g(\mathbf{x}(t))=\tau f(\mathbf{x})+\mathbf{x}$
and we scale the input $\mathbf{u}(t)$ by $\tau$. For arbitrary
dynamics of the form \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}f(\mathbf{x},\mathbf{u})$},
we do a change of variables $\mathbf{x}^{\prime}=\langle\mathbf{x},\mathbf{u}\rangle$.

This exploitation of the inherent first-order low-pass filter found
in synaptic connections allows for the implementation of a very wide
variety of systems, including oscillators, integrators, and arbitrary
attractor networks \cite{Eliasmith2005p}. 
This approach allows for the construction of neural models that correspond
to a very large family of functions, including those typically employed
by modern control theory and dynamic systems theory.

@@

The principles of the NEF have been incorporated into the Nengo neural simulator \cite{Bekolay2014}.  Nengo is a general purpose simulator that incorporates the three principles of the NEF.  It has been used to build what is currently the world's largest functional brain model \cite{Eliasmith2012b}, along with many other simulations. Full details and documentation can be found online at \url{http://github.com/nengo/nengo}.

One central feature of Nengo is that it permits the same model description to target a variety of ``backends.''  Current backends include CPUs, GPUs, the IBM BlueGene supercomputer, and several neuromorphic chips, including SpiNNaker \cite{Khan2008} and Neurogrid \cite{Choudhary2012}.  Because Nengo exploits the NEF methods, model descriptions can be very concise.  Model descriptions are written in python, although backends compile to a variety of languages, depending on the target hardware. For example, to create a recurrent model of a chaotic dynamical system (a Lorenz attractor) with 2000 spiking neurons and run it for 10 seconds, the following suffices:

\begin{verbatim}
import nengo

model = nengo.Network()
with model:
    state = nengo.Ensemble(2000, 3, radius=60)

	def feedback(x):
	    dx0 = -10 * x[0] + 10 * x[1]
	    dx1 = -x[0] * x[2] - x[1]
	    dx2 = x[0] * x[1] - 8.0/3 * (x[2] + 28) - 28

	    return [dx0 * 0.1 + x[0],
	            dx1 * 0.1 + x[1],
	            dx2 * 0.1 + x[2]]

    nengo.Connection(state, state, function=feedback, synapse=0.1)
    state_probe = nengo.Probe(state, synapse=tau)

sim = nengo.Simulator(model)
sim.run(10)
\end{verbatim}

Nengo will use the NEF methods to automatically solve for the connection weights that will
best approximate the provided nonlinear, dynamic function.  To run this same network on a different backend, we only need change the second last line to indicate which simulator to use.

Nengo also provides an interactive interface for building models, and displaying the results of a simulation while it is running, allowing for real-time interaction
with a running model.

In this paper, we describe an extension to the Nengo simulator that allows models to compile to an FPGA, using a simulation method that efficiently approximates the computations performed by populations of spiking neurons.

@@

- Bryan???
- mention the other modes first to provide context/more info about Nengo

@@

\design{} is a custom digital hardware design that implements the population-mode NEF in real-time (1~ms timestep) on an FPGA.
A block diagram of the design is shown in Figure~\ref{fig:system}.
The encoder units perform the function of encoding (), % FIXME NEF equation
while the population units perform the decoding () % FIXME NEF equation
and write the decoded values to the decoded value buffers
to be used as inputs to other populations or as outputs from the network.
Each population of neurons is assigned to a population unit, which is able to simulate up to 1024 populations
having either one or two encoded inputs each. These inputs are calculated by the encoder units,
of which there are two for a one-dimensional population unit and four for a two-dimensional population unit;
each encoder unit can be associated with a different first-order filter to simulate post-synaptic filtering.
For example, one of the two encoders on a one-dimensional population unit can be programmed with
a small time constant to model fast synapses (e.g. AMPA and GABA$_A$), while the second can be programmed
with a large time constant to model slow synapses (e.g. NMDA and GABA$_B$).
Each population unit performs time-multiplexing and swaps in different sets of decoders
in order to run more populations on the same hardware. The principal components are fixed and must be shared
between all populations on an individual population unit.
Each population unit writes to a fixed set of decoded value buffers, but a shared interconnect
allows the encoder units to perform random accesses to these buffers.
This allows effective all-to-all communication between populations while reducing the number of 
connections that must be made between population units.
A number of decoded value buffers are reserved for external inputs to the network,
which can be sent from the controlling computer system over an Ethernet connection.
The output channels can also be programmed to send decoded values back to the computer over Ethernet.

@@

Nengo supports a more advanced connection called a learning connection,
which modifies the decoders of a population according to a particular ``learning rule''
as a function of other represented values. The hardware design does not support this special connection type.
It was decided to implement the most common cases first in order to demonstrate the feasibility of simulating
neural networks in population mode on the hardware, and since many networks do not use learning connections,
support for them was omitted from this version.
